The type graph method is a technique for proving termination of DPO rewriting systems. 
To apply this method, one needs to construct a weighted type graph over a semiring that witnesses termination of the rewriting system, and three concrete well-founded semirings over the natural numbers have been proposed in prior work: the tropical semiring, the arctic semiring, and the arithmetic semiring.
However, constructing suitable weighted type graphs over these concrete semirings for DPO rewriting systems on edge-labeled directed multigraphs is difficult. 
To address this challenge, we investigate weighted type graphs over non-well-founded semirings, and propose three corresponding concrete semirings over the real numbers: the tropical semiring, the arctic semiring, and the arithmetic semiring. Construction of weighted type graphs over these concrete semirings is computationally easier.
    
We implemented both our approach and the one proposed by Endrullis et al.~\cite{endrullis2024generalized_arxiv_v2} for DPO rewriting systems on edge-labeled directed multigraphs into a unified tool. Experiments show that for the tropical and arctic semirings, the weighted type graphs over the real semiring can be constructed more easily, but for the arithmetic semiring, the weighted type graphs over the real semiring are more difficult to construct than those over the natural semiring.
% , because the complexity is extremely high and the search space cannot be restricted by bounding the maximum edge weight. 

A promising direction involves generalizing the type graph method to other algebraic graph rewriting formalisms, broadening its applicability in automated verification. 

\section*{Acknowledgments}
I thank the anonymous reviewers for their valuable feedback; my supervisor, Xavier Urbain, for his guidance; and my mother and wife for their unwavering support.
% \begin{credits}
%   \ifanonymous   
%   \else
%   \trackedtext{
%       \subsubsection{\ackname} 
%       I thank my supervisor, Xavier Urbain, for his guidance; the anonymous reviewers for their valuable feedback; Rachid Echahed for proofreading; and my mother and wife for their unwavering support.
%       % This study was partially funded by the project SAPPORO, ANR, 2019-CE25-0005.
%       \subsubsection{\discintname}
%       The author have no competing interests to declare that are relevant to the content of this article.
%     }
%     \fi 
% \end{credits} 