First of all, I thank my supervisor, Xavier Urbain, for providing me with the opportunity to work on this subject and for his guidance not only in my research but, more importantly, in my personal development: his integrity is a great example for me to follow. Midway through my PhD, we had different opinions regarding the research direction, and he generously allowed me to pursue the path which inspired me the most. This freedom has been invaluable to me, and I am grateful for his kindness. I am also grateful for his careful, tireless reading of this dissertation; when I initially disagreed with his suggestions he was patient and later restated them in a way I found considerate. His help significantly improved the quality of my writing.

I am grateful to the members of the jury: Clara Bertolissi, Évelyne Contejean, Pierre Hyvernat, Jean Krivine, Pascale Le Gall, Philippe Malbos and Frédéric Prost, for their time, careful reading and constructive comments.

I thank Evelyne Contejean and Victor Ostromoukhov for serving as members of the \enquote{Comité de Suivi de Thèse}; Pierre Courtieu for his help in the early stages of my PhD and for his extraordinary kindness; Rachid Echahed, Nicolas Peltier, Sophie Tourret and Stephan Merz for hosting me as temporary teaching and research associate (ATER) at the University of Grenoble Alpes and at the University of Lorraine; Tom Hirschowitz for his invitation to give a talk at the University of Chambéry, which meant a lot to me as a follower of Jean-Jacques Rousseau; 
Zhilin Wu for his constructive questions on the first draft;
the anonymous reviewers of my work for their time and valuable feedback;
Benoît Mariou, whose course first inspired me to pursue logic; Jean-marie Lagniez for his encouragements during this thesis.

I extend my sincere thanks to all the faculty members who contributed to my academic development during my studies in France, for their generosity in sharing knowledge, and for their kindness and patience towards their students. I am especially grateful to Benoît Mariou, who taught my first course in logic; it was in that class that I realized logic was what I had been looking for.

I thank my parents, Shuangluan Ye and Guowen Qiu, for their unconditional love and sacrifices that have allowed me to pursue my dreams. I also thank my wife, Lu Wang, for her patience and understanding during the challenging times of my PhD.

I thank my friend Pierre Bourquat, the most authentic person I have ever met. I still remember one day at the very beginning of my PhD, when I was working in my cubicle alone, he gently tapped on the glass side of my cubicle and asked me if I was new ... then introduced me to other friends Thomas Ranvier, Auday Berro, Mehdi Hennequin and Yacine Gaci. I also thank Zhongyun Yang, Yaqing Fang, Jiayi Wei, Chukun Zhang, Xin Xie and all those who have come into my life, as each person has played a role in shaping who I am today.

The author acknowledges the use of large language models, but solely for the purpose of proofreading and wording edits.

Finally, I would like to express my gratitude to the French National Research Agency (ANR) for funding the project Safe, Adaptive, and Provable Protocols for Oblivious Robots Operation (SAPPORO) under Grant 2019-CE25-0005, which has provided essential resources and support for my research endeavors.  
