We define binary relations and orders on a collection of objects instead of on a set. This is because later we will define rewriting relations as binary relations, and the collection of objects subject to be rewritten in a rewriting system is not necessarily a set. For example, the collection of finite directed edge-labeled graphs is not a set, but a class.

\begin{notation}
    Let $C$ and $C'$ be collections of objects. We denote $c \mathop{\in} C$ for~\enquote{$c$ is an element of C}, $C \mathop{\cup} C'$ the smallest collection contains all elements in $C$ and all elements in $C'$, $C \mathop{\cap} C'$ the smallest collection contains all elements in $C$ that are also in $C'$, $C * C'$ the collection of ordered pairs with first element in $C$ and second element in $C'$.  
  \end{notation} 
  
  \begin{definition}[Binary relation]
    \label{def:binary_relation:binary_relation}
    Let $D$ be a collection of objects. A mathematical structure \( (D, \mathcal{R}) \) where $\mathcal{R}$ is a collection of objects from $D * D$ is called a \textbf{binary relation} on $D$. 
    
    For an object in $\mathcal{R}$ with first element $x \mathop{\in} D$ and second element $y\in D$, we write $x \mathcal{R} y$, $\mathcal{R}(x,y)$ or $(x,y) \mathop{\in} \mathcal{R}$. 
    
    When $D$ is irrelevant, we say simply that $\mathcal{R}$ is a binary relation.
  \end{definition} 
   
  \begin{definition}[Reflexivity,
    %  Antisymmetry, 
     Transitivity]
    \label{def:binary_relation:reflexivity_transitivity}
    A binary relation \( \mathop{\to} \) on \(C\) is said to be \textbf{reflexive} if for every object \(a \mathop{\in} C\), we have \(a \mathop{\to} a\); \textbf{transitive} if for all objects \( a, b, c \mathop{\in} C\), \( a \mathop{\to} b \mathop{\land} (b \mathop{\to} c) \) implies \(a \mathop{\to} c\).
  %       \item \textbf{antisymmetric} if \(
  % \forall a, b \mathop{\in} A, \, (a \mathop{\to} b) \mathop{\land} (b \mathop{\to} a) \implies a \mathop{=} b.
  % \)
  \end{definition}
  
  \begin{definition}[$\mathcal{R}$-sequence]
    \label{def:binary_relation:sequence}
    Let \(\mathcal{R}\) be a binary relation on $S$.
    A \textbf{\( \mathcal{R} \)-sequence} is either a finite sequence \( \left( s_i \right)_{0 \leq i \leq m} \) of elements in $S$ such that \(s_i \mathcal{R} s_{i+1}\) for each \( 0 \leq i \leq m-1\), or an infinite sequence \((s_i)_{i \mathop{\in} \mathbb{N}}\) of elements in $S$ such that \(s_n \mathcal{R} s_{n+1}\) for each \(i \mathop{\in} \mathbb{N}\).
\end{definition}

\begin{definition}[Well-founded binary relation]
    \label{def:binary_relation:well_founded}
    A binary relation $\to$ on a collection $S$ of object is said to be \textbf{well-founded} if there is no infinite $\to$-sequence. 
    % In other words, there does not exist an infinite sequence of elements $ (s_i)_{i \mathop{\in} \mathbb{N}} $ where $s_i \mathop{\in} S $ such that $s_i \mathop{\to} s_{i+1}$ for $i \mathop{\in} \mathbb{N}$.
\end{definition}

\begin{definition}[Transitive closure]
    \label{def:binary_relation:transitive_closure}
    The \textbf{transitive closure} of a binary relation $\to$, denoted $\to^+$, is the smallest transitive relation that include \( \mathop{\to} \).
  \end{definition}
  
  \begin{definition}[Reflexive-transitive closure]
    \label{def:binary_relation:reflexive_transitive_closure}
    The \textbf{reflexive transitive closure} of a binary relation $\to$, denoted $\to^*$, is the smallest reflexive and transitive relation that include \( \mathop{\to} \).
  \end{definition}

  \begin{definition}[Homomorphisms of binary relations]
    \label{def:binary_relation:homomorphism}
    Let $\to_\mathcal{A}$ be a binary relation on a collection $A$ of objects and $\to_\mathcal{B}$ be a binary relation on a collection $B$ of objects. A \textbf{homomorphism} from $\to_\mathcal{A}$ to $\to_\mathcal{B}$ is a function \( h: A \mathop{\to} B \) such that for all \( a, b \mathop{\in} A \), if \( a \mathop{\to} _\mathcal{A} b \) then \( h(a) \mathop{\to} _\mathcal{B} h(b) \).
  \end{definition}
  
\begin{proposition}[Proving well-foundedness]
  \label{prop:binary_relation:proving_well_foundedness}
  Let \(\to\) be binary relation and let $\leadsto$ be a well-founded binary relation. If there is a homomorphism from \(\to\) to \(\leadsto\), then \(\to\) is well-founded.
\end{proposition}